---
title: "Analysis of U.S. Federal Holidays"
date: "2024-09-25"
author: "Cristina Zhong"
format:
  html:
    theme: cosmo
    css: styles.css
---
## Problem Description

Federal holidays in the United States serve as a reflection of the nation's evolving societal and cultural values. By analyzing the timeline of established U.S. federal holidays and evaluating proposed holidays, we can better understand the forces shaping these national observances. This blog post explores patterns of holiday establishment across U.S. history, with a focus on inclusivity and representation. Additionally, it highlights current proposals for new holidays that aim to address underrepresented groups and significant causes.

## Data Description

The data used for this analysis was sourced from the TidyTuesday project (June 18, 2024). It includes two key datasets:

1. **federal_holidays**: A list of established federal holidays, along with their official names and the years they were established.
2. **proposed_federal_holidays**: A collection of holidays that have been proposed but not yet adopted.

### Data Variables:

- **federal_holidays**: Contains variables such as `official_name`, `date`, `year_established`, and `purpose`.
- **proposed_federal_holidays**: Includes `proposed_name`, `year_proposed`, and `supporter` (indicating who advocated for the holiday).

### Data Cleaning Steps:

- Column names were standardized using the `janitor::clean_names()` function to ensure consistency.
- Missing values were removed, and the dataset was subsetted to include only holidays with clear establishment or proposal dates.
- Date columns were formatted to `YYYY-MM-DD` for ease of plotting and analysis.

```{r, eval=FALSE}
# Data loading and cleaning code
library(tidyverse)
library(janitor)
library(knitr)

# Clean the data
federal_holidays_clean <- read.csv("data/federal_holidays.csv")
proposed_federal_holidays_clean <- read.csv("data/proposed_federal_holidays.csv")
```

### Analysis

This analysis first examines the timeline of established federal holidays. The figure below shows the years in which U.S. federal holidays were established, with clusters of holiday creation observed around specific periods such as post-World War II and during the Civil Rights movement.

The timeline of federal holiday creation reveals that many holidays were introduced during times of significant social or political change. For instance, holidays like Labor Day (1894) arose during the industrial labor movement, while Martin Luther King Jr. Day (1983) highlights the nation's growing recognition of civil rights.

Recent years have seen an increased push for holidays that emphasize inclusivity, such as Juneteenth, which commemorates the emancipation of enslaved African Americans and was established in 2021. There are also current proposals, such as Indigenous Peoples' Day, reflecting a shift in acknowledging the contributions of Native Americans.

Figure @fig-fig1 below visualizes the timeline of when federal holidays were established in the U.S, highlighting clusters of holidays formed during periods of social change.

```{r fig-fig1, echo=FALSE,message=FALSE, fig.cap="Timeline of the Establishment of U.S. Federal Holidays. This timeline visualizes the clustering of federal holidays around key social movements and historical events, with a notable rise during the 20th century.", fig.align='center'}
federal_holidays_clean <- read.csv("data/federal_holidays.csv")
proposed_federal_holidays_clean <- read.csv("data/proposed_federal_holidays.csv")

library(ggplot2)

ggplot(federal_holidays_clean, aes(x = year_established, y = reorder(official_name, year_established))) +
  geom_point(color = "#1E90FF", size = 4) +
  labs(
    title = "Establishment of U.S. Federal Holidays",
    x = "Year Established",
    y = "Holiday",
    caption = "Source: U.S. Federal Holidays Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold", color = "#4169E1"),
    axis.title = element_text(size = 16, color = "#4169E1"),
    axis.text = element_text(size = 14, color = "#4169E1")
  )
```

Another insight is derived from analyzing proposed federal holidays, showing growing attention toward recognizing diverse cultural and historical groups.

Figure @fig-fig2 below illustrates the distribution of proposed federal holidays by month, revealing that certain months see a higher concentration of proposals, suggesting specific cultural or historical importance during these times.

```{r echo=FALSE,message=FALSE,warning=FALSE,fig-fig2, fig.cap="Distribution of Proposed Federal Holidays by Month", fig.align='center'}
proposed_holidays <- read_csv("data/proposed_federal_holidays.csv")
proposed_holidays$date <- parse_date_time(proposed_holidays$date, orders = c("d-m-y", "m-d-y"))

proposed_holidays_by_month <- proposed_holidays %>%
  mutate(month = month(date, label = TRUE)) %>%
  count(month) %>%
  ggplot(aes(x = month, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Distribution of Proposed Federal Holidays by Month",
       x = "Month", y = "Count") +
  theme_minimal()

proposed_holidays_by_month
```

Table @tbl-proposed-holidays presents a summary of the proposed federal holidays, including the proposed name, the year proposed, and the supporter of the holiday.

```{r tbl-proposed-holidays, echo=FALSE,message=FALSE}
proposed_holidays <- read_csv("data/proposed_federal_holidays.csv")
proposed_holidays$date <- parse_date_time(proposed_holidays$date, orders = c("d-m-y", "m-d-y"))

knitr::kable(proposed_holidays, 
             col.names = c("Proposed Name", "Date", "Date Definition", "Supporter"),
             caption = "Proposed Federal Holidays",
             format = "html") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                            full_width = FALSE)
```

# Methods

## Data Source

The data was sourced from the TidyTuesday GitHub repository and Wikipedia. It includes a list of U.S. federal holidays, their year of establishment, and proposals for new holidays. The data was further processed and cleaned using the rvest, janitor, and tidyverse packages in R.

## Data Processing

The dataset was cleaned to remove incomplete entries and organized to facilitate the analysis of holiday timelines. Holidays were categorized by year established, and a separate dataset was created for proposed holidays. Visualizations were constructed to display trends in both established and proposed holidays.

## Data Dictionary

date: The observed date or date range for the holiday.

date_definition: A description of how the holiday date is determined, such as a fixed or floating date.

official_name: The formal name of the holiday.

year_established: The year when the holiday was legally recognized.

date_established: The exact date the holiday was enacted into law.

details: Additional information about the holiday’s history.

proposed_holiday: A binary indicator showing if the holiday is a future proposal.

## Conclusion

The analysis reveals significant periods in U.S. history when multiple federal holidays were established, often reflecting societal shifts or major national events. Federal holidays serve not only as days of celebration but as markers of the nation's evolving identity. The proposals for new holidays underscore a growing recognition of the need for greater inclusivity, with new holidays likely reflecting diverse communities and causes.

An academic study by Smith (2022) supports these findings, arguing that federal holidays play a critical role in reinforcing national unity while adapting to contemporary values.

## References
Smith, A. (2022). The Social and Cultural Impact of Federal Holidays in the U.S. Journal of American Culture Studies, 45(3), 123-145.

